---
title: "Compassionate Release Data Analysis"
author: "Hope Johnson"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(readxl)
library(skimr)
library(kableExtra)
library(behindbarstools)
library(rmarkdown)
library(glue)
library(lubridate)

knitr::opts_chunk$set(fig.height=13) 
knitr::opts_chunk$set(message = FALSE)
dat_path <- here("compassionate-release-denials", "data")
plot_path <- here("compassionate-release-denials", "plots")
```

## Read clean data 

Read in the most up-to-date data that Rebecca and volunteers have cleaned:

```{r, warning=FALSE, message=FALSE}
ucla_dat <- read_csv(file.path(dat_path, "2021-06-11-master.csv")) %>%
  ## filter out duplicates marked by Rebecca
  filter(is.na(`4-8 dupes`)) %>%
  mutate(merging_ucla = glue('{docketno_std}_{court_std}')) %>%
  select_if(~sum(!is.na(.)) > 0) %>% # rm all NA
  mutate(date_filed_std = as.Date(date_filed_std),
         court_std = ifelse(court_std == "E.D. Louisiana.", "E.D. Louisiana", court_std)) 

nrow(ucla_dat)
```

Merge Tori Finkle's data set with the UCLA data set:

```{r, warning=FALSE, message=FALSE}
finkle_data <- read_xlsx(file.path(dat_path, "2020-02-16-cr_finkle.xlsx")) %>%
  mutate(court_std = ifelse(court_std == "D. Hawai'i", "D. Hawaii", court_std), 
         merging_finkle = glue('{docketno_std}_{court_std}')) %>%
  select_if(~sum(!is.na(.)) > 0) %>% # rm all NA
  mutate(Date_finkle = ifelse(!is.na(Date), glue('{substr(Date, 5, 7)} {substr(Date, 9, 10)} {substr(Date, 25, 28)}'), NA)) %>%
  mutate(Date_finkle_parsed = mdy(Date_finkle))

merged <- full_join(ucla_dat, finkle_data, by = c("merging_ucla" = "merging_finkle",
                                                  "def_last_name" = "defendant_last_name"))
```

```{r, warning=FALSE, message=FALSE}
dat_merged <- merged %>% 
  mutate(finkle_lower = tolower(Decision)) %>%
  mutate(court_std = behindbarstools::coalesce_with_warnings(court_std.x, court_std.y)) %>% 
  mutate(docketno_std = behindbarstools::coalesce_with_warnings(docketno_std.x, docketno_std.y)) %>% 
  ## create bool for times that ucla and finkle classifications were both non-NA and different 
  mutate(diff_status = ifelse(!is.na(finkle_lower) & !is.na(granted_denied) & (finkle_lower != granted_denied),
                              TRUE, FALSE)) %>%
  ## create bool for finkle more recent than ucla (when both non-NA and different)
  mutate(status_mostrecent = ifelse(diff_status, Date_finkle_parsed > date_filed_std, NA)) %>%
  mutate(status_alldiff = behindbarstools::coalesce_with_warnings(granted_denied, finkle_lower)) %>%
  ## first, grab the first non-NA date
  mutate(date_coalesced_first = coalesce(Date_finkle_parsed, date_filed_std),
         date_both = ifelse((!is.na(Date_finkle_parsed) & !is.na(date_filed_std)), TRUE, FALSE)) %>%
  ## then create a condition when both dates are present -- grab the latest one
  mutate(date_coalesced = case_when(
    date_both == TRUE & Date_finkle_parsed > date_filed_std ~ Date_finkle_parsed,
    date_both == TRUE & date_filed_std > Date_finkle_parsed ~ date_filed_std,
    date_both == TRUE & date_filed_std == Date_finkle_parsed ~ date_filed_std,
    date_both == FALSE ~ date_coalesced_first)
    ) %>%
  mutate(judge_coalesced = behindbarstools::coalesce_with_warnings(judge_descr, Judge),
         # def_last_name_coalesced = behindbarstools::coalesce_with_warnings(defendant_last_name, def_last_name),
         case_name_coalesced = behindbarstools::coalesce_with_warnings(`Case Name`, Case)) %>%
  ## create `decision` based off of most recent case when diverging classification, and coalesced version otherwise
  mutate(status = case_when(
    status_mostrecent == TRUE ~ finkle_lower,
    status_mostrecent == FALSE  ~ granted_denied,
    is.na(status_mostrecent) ~ status_alldiff)) %>%
  dplyr::select(case_name_coalesced,
         docketno_std,
         date_coalesced,
         court_std,
         judge_coalesced,
         status,
         `Document Description`,
         Title,
         def_last_name,
         `Nature of Suit`,
         `Prosecutor Name`,
         ucla_citation = Citation.x,
         finkle_Cite = Citation.y,
         URL,
         Link, ## coalesce this with "URL"? 
         `Parallel Cite`
         )

head(dat_merged)

# ## TO DO : grab status from whoever has the most recently updated document
# ##### there are 168 of these situations
# diff_status <- dat %>%
#   filter(!is.na(finkle_lower) & !is.na(granted_denied) & (finkle_lower != granted_denied)) %>%
#   mutate(more_recent = ifelse(Date_finkle_parsed > date_filed_std, "finkle", "ucla")) %>%
#   select(docket_no_std, finkle_lower, granted_denied, date_filed_std, Date_finkle_parsed, more_recent)
# nrow(diff_status)

```

Import the yearly criminal case by court data, merge with our data, format all the data together for mapping.

Data source for total number of criminal cases sourced here: https://www.uscourts.gov/statistics/table/d-3/statistical-tables-federal-judiciary/2020/12/31 

Map of federal courts here: https://hifld-geoplatform.opendata.arcgis.com/datasets/us-district-court-jurisdictions 

First, prep the data by merging all three sources: 

```{r}
## read in federal court crosswalk
court_xwalk <- read_csv(file.path("~", "UCLA", "misc-data", "federal courts", "courts_mapping.csv"))

## read in yearly criminal case data 
crim_dat_path <- file.path(dat_path, "stfj_d3_1231.2020.xlsx")
crim_dat <- read_xlsx(crim_dat_path, skip = 2)
crim_dat_cln <- crim_dat %>%
  select_if(~sum(!is.na(.)) > 0) %>% # rm 100% missing cols 
  mutate(across(where(is.character), ~na_if(., "-"))) 
names(crim_dat_cln) <- as_vector(ifelse(!is.na(crim_dat_cln[1, ]), 
                                        crim_dat_cln[1, ], 
                                        names(crim_dat_cln)))

## read in geospatial data 
## District of Columbia > anomaly
map_dat <- sf::read_sf(file.path("~", "UCLA", "misc-data", "federal courts", "US_District_Court_Jurisdictions.json")) %>%
  dplyr::select(STATEFP, NAME, STATE, CHIEF_JUDG, NOMINATING, TERM_AS_CH, DISTRICT_N, geometry)
split <- str_split_fixed(map_dat$NAME, "District of ", n = 2)
dist <- split[,1]
state <- split[,2]
dist_trans <- ifelse(dist != "", 
                     glue("{str_sub(dist, 1, 1)}.D."),
                     "D.")
map_dat_cln <- map_dat %>% 
  mutate(court_standardized = ifelse(state == "District of Columbia", 
                                 state,
                                 glue("{dist_trans} {state}")),
         court_standardized = ifelse(court_standardized == "D. Northern Marianas Islands",
                                     "D. Northern Mariana Islands",
                                     court_standardized),
        court_standardized = ifelse(court_standardized == "D. US Virgin Islands",
                           "D. Virgin Islands",
                           court_standardized)
  )
         
## prep yearly criminal case data for merging
crim_dat_out <- crim_dat_cln %>%
  filter(!is.na(`Circuit and District`)) %>%
  left_join(court_xwalk, by = c("Circuit and District" = "Court statistics")) %>%
  filter(!is.na(District)) %>% 
  dplyr::select(District, Total) %>% ## only keep totals for now 
  rename(all_crim_cases = Total)

## merge yearly crim case data by court with CR, geospatial 
islands <- c("Hawaii", "Alaska", "US Virgin Islands", 
             "Northern Marianas Islands", "Puerto Rico")

dat <- dat_merged %>%
  # tab_2 %>%
  left_join(crim_dat_out, by = c("court_std" = "District")) %>%
  left_join(map_dat_cln, by = c("court_std" = "court_standardized")) %>%
  mutate(continental_us = ifelse(STATE %in% islands, FALSE, TRUE)) %>%
  mutate(facet_group = ifelse(continental_us, "Continental US", STATE),
         finkle_Cite = ifelse(str_detect(finkle_Cite, "(?i)slip copy"), NA, finkle_Cite)) %>%
  mutate(district_name_ussc = str_remove(NAME, "District of ")) %>%
  rename(case_name = case_name_coalesced,
       docket_number = docketno_std,
       date = date_coalesced,
       court_std = court_std,
       judge = judge_coalesced,
       granted_denied = status,
       doc_description = `Document Description`,
       doc_title = Title,
       def_last_name = def_last_name,
       nature_of_suit = `Nature of Suit`,
       prosecutor_name = `Prosecutor Name`,
       published_cite = finkle_Cite,
       courtlistener_url = URL,
       westlaw_url = Link,
       westlaw_cite = `Parallel Cite`,
       state_fips = STATEFP,
       state = STATE,
       full_court_name = NAME,
       chief_judge = CHIEF_JUDG,
       term_as_ch = TERM_AS_CH,
       nominating = NOMINATING,
       district_n = DISTRICT_N,
       district_name_ussc = district_name_ussc)

dat_out <- dat %>%
  dplyr::select(case_name,
         docket_number,
         date,
         court_std,
         judge,
         granted_denied,
         doc_description,
         doc_title,
         def_last_name,
         nature_of_suit,
         prosecutor_name,
         ucla_citation,
         published_cite,
         courtlistener_url,
         westlaw_url,
         westlaw_cite,
         state_fips,
         state,
         full_court_name,
         chief_judge,
         nominating,
         term_as_ch,
         district_n)
#write_csv(dat_out, file.path(dat_path, "CR_data.csv"), na = "")
```


## Summarize data

The data spans from `r min(dat$date)` to `r max(dat$date)`. This data set includes `r nrow(dat)` distinct documents, and `r n_distinct(dat$docketno_std)` distinct cases. 

```{r}
skim(dat)
# n_distinct(dat$`Docket Number`) 
n_distinct(dat$docket_number)
```
## Add region indicator

We use state region classifications from US Census Bureau Regions and Divisions by State: http://www.census.gov/geo/maps-data/maps/pdfs/reference/us_regdiv.pdf. Thanks https://github.com/cphalpert/ for making it into a CSV!

```{r}
regions <- read_csv("https://raw.githubusercontent.com/cphalpert/census-regions/master/us%20census%20bureau%20regions%20and%20divisions.csv")

dat <- dat %>%
  mutate(court_state = gsub('.*D. (.*)', '\\1', court_std)) %>%
  left_join(regions, by = c("court_state" = "State")) 
```

## Plots, analysis

First, we calculate the percentage granted and denied. 

```{r}
tab_1 <- dat %>% 
  mutate(document_count = nrow(dat)) %>%
  group_by(granted_denied) %>%
  summarise(
            group_count = n(),
            n_documents = first(document_count)) %>%
  ungroup() %>%
  mutate(ntl_percent = (group_count / n_documents) * 100)

tab_1 %>%
  mutate(ntl_percent_rounded = round(ntl_percent, digits = 1)) %>% 
  dplyr::select(Status = granted_denied, 
       Count = group_count,
       Percentage = ntl_percent_rounded) %>%
  paged_table()
```

How many documents do we have by court? 

```{r, echo = FALSE}
all_docs <- nrow(dat)
group_counts <- dat %>%
  group_by(court_std) %>%
  summarise(n_documents_by_court = n()) %>%
  ungroup() %>%
  mutate(perc_total = (n_documents_by_court / all_docs) * 100) %>%
  arrange(desc(n_documents_by_court)) 

group_counts %>%
  mutate(perc_total_rounded = round(perc_total, 2)) %>%
  dplyr::select(`Court Name` = court_std,
         `Document Count` = n_documents_by_court,
         `Percent of Total` = perc_total_rounded) %>%
  paged_table()
  
# make a plot out of it!
plt <- ggplot(group_counts, 
       aes(y = n_documents_by_court, 
           x = reorder(court_std, n_documents_by_court))) + 
  geom_bar(position="dodge", stat="identity") +
  coord_flip() + 
  theme_minimal() + 
  labs(x = "", y = "# Documents by Court") 
# plt
```

We're interested in the percentage granted and denied broken down by court. 

```{r}
# rename tab 1 for merge-ability 
tab_1_merging <- tab_1 %>%
  dplyr::select(granted_denied,
         ntl_percent)

tab_2 <- dat %>% 
  group_by(granted_denied, court_std) %>%
  summarise(group_count = n(),
            Region = first(Region),
            Division = first(Division)) %>%
  ungroup() %>%
  left_join(group_counts) %>%
  mutate(court_percent = (group_count / n_documents_by_court) * 100) %>%
  # merge in national granted/denied breakdown
  left_join(tab_1_merging, by = "granted_denied") %>%
  mutate(diff = ntl_percent - court_percent) %>%
  arrange(desc(n_documents_by_court), desc(abs(diff)))

tab_2 %>%
  dplyr::select(-c(court_percent, ntl_percent, diff)) %>%
  pivot_wider(names_from = granted_denied,
              values_from = group_count) %>%
  mutate(pct_denied = (denied / n_documents_by_court) * 100,
         pct_granted = (granted / n_documents_by_court) * 100,
         pct_denied_rounded = round(pct_denied, 1),
         pct_granted_rounded = round(pct_granted, 1),
         grants = glue('{granted} ({pct_granted_rounded}%)'),
         denials = glue('{denied} ({pct_denied_rounded}%)')) %>%
  arrange(-n_documents_by_court) %>% 
  dplyr::select(`Court Name` = court_std,
         `Document Count` = n_documents_by_court,
         `Denied` = denials,
         `Granted` = grants) %>%
  paged_table()
```

Were there any courts with significantly more or less compassionate release cases granted than the national average? 

```{r}
ntl_pct_granted <- tab_1_merging %>% 
  filter(granted_denied == "granted") %>%
  pull(ntl_percent)

# filter out courts that had less than 15 total CR documents
court_specific_grants <- tab_2 %>%
  filter(granted_denied == "granted",
         n_documents_by_court > 14) %>%
  ggplot(., 
       aes(y = court_percent, 
           x = reorder(court_std, diff),
           color= Region,
           fill = Region)) + 
  geom_bar(position="dodge", stat="identity") +
  geom_hline(yintercept = ntl_pct_granted, 
             linetype="dotted", 
                size=1.5) + 
  labs(x = "", y = "% Granted",
       subtitle = "Black dotted line shows national average % granted") + 
  coord_flip() + 
  theme_minimal()
ggsave(file.path(plot_path, "court_specific_grants_region.png"), court_specific_grants, height = 10, width = 14)

```

It appears that most of the courts with lower than average grants were located in the south. Mapping it will be better for understanding. 

Now that the data is ready, map it: 

```{r}
## map (# CR) / (total criminal cases)
small_name_xwalk <- dat %>%
  dplyr::select(court_std, district_name_ussc) %>%
  unique()

mapping_dat_out <- dat %>% 
  group_by(granted_denied, court_std,.drop = FALSE) %>%
  summarise(group_count = n(),
            all_crim_cases = first(all_crim_cases)) %>% ## calc # granted, # denied
  ungroup() %>%
  left_join(group_counts, by = "court_std") %>% 
  dplyr::select(-perc_total) %>% 
  pivot_wider(names_from = "granted_denied", values_from = "group_count") %>% 
  mutate(perc_granted = granted / n_documents_by_court,
         perc_denied = denied / n_documents_by_court,
         perc_cr = n_documents_by_court / all_crim_cases) %>%
  right_join(map_dat_cln, by = c("court_std" = "court_standardized")) %>%
  left_join(small_name_xwalk, by = "court_std") %>%
  left_join(ussc, by = c("district_name_ussc" = "district")) %>%
  mutate(diff_ussc = n_documents_by_court - ussc_total,
         diff_granted = granted - ussc_n_granted,
         diff_denied = denied - ussc_n_denied,
         perc_denied_ussc = ussc_n_denied / ussc_total)
st_write(mapping_dat_out, file.path(dat_path, "court_mapping_data.shp"),
         append = FALSE)

## decided to map it in QGIS 
# map_1 <- dat %>%
#   group_by(court_std) %>%
#   filter(row_number() == 1) %>% 
#   ungroup() %>%
#   mutate(prop_cr = n_documents_by_court / all_crim_cases) 
# 
# map_1 %>%
#   filter(continental_us == TRUE) %>%
#   mutate(prop_cr_cut = cut_interval(prop_cr, n = 6)) %>%
#   ggplot(data = ., 
#        aes(geometry = geometry)) + 
#   geom_sf(aes(fill = prop_cr_cut)) + 
#     scale_fill_discrete(
#     name = "Proportion of CR cases",
#     labels = c("0 - 14%", 
#                "14 - 28%", 
#                "28 - 42%",
#                "42 - 56%",
#                "56 - 70%",
#                "70 - 84%")
#   ) + 
#   # dont think this is changing anything, 
#   coord_sf(
#     xlim = c(-20000000, -3000000),
#     ylim = c(3000000, 11000000)
#   ) +
#   ggthemes::theme_map() #+ 
#   # facet_wrap(~facet_group)
# 
# ## map % granted 
# map_1 %>%
#   filter(continental_us == TRUE,
#          granted_denied == "granted") %>%
#   mutate(court_percent_cut = cut_interval(court_percent, n = 3)) %>%
#   ggplot(., 
#        aes(geometry = geometry)) + 
#   geom_sf(aes(fill = court_percent_cut)) + 
#   coord_sf(
#     xlim = c(-20000000, -5000000),
#     ylim = c(2000000, 12000000)
#   ) +
#   ggthemes::theme_map()

```
## Comparison to U.S. Sentencing Commission data

Source: <https://www.ussc.gov/sites/default/files/pdf/research-and-publications/federal-sentencing-statistics/compassionate-release/20210609-Compassionate-Release.pdf>, data extracted from table 1. 

```{r}
ussc <- read_csv(file.path(dat_path, "ussc_cr.csv")) %>%
  dplyr::select(district = district,
         ussc_total = total,
         ussc_n_granted = n_granted,
         ussc_n_denied = n_denied)
```



### Time series

Did the number of compassionate release cases change over the course of the pandemic? 

```{r}
## all - weekly
timebars_all_weekly <- dat %>%
  group_by(week = week(date), year = year(date), granted_denied) %>% 
  summarize(count = n()) %>%
  ungroup() %>%
  # mutate(week_date = as.Date(glue('{year}-{week}-1'), "%Y-%U-%u")) %>%
  mutate(week_date = lubridate::ymd( glue("{year}-01-01" )) + lubridate::weeks(week - 1 )) %>%
  ggplot(., aes(x = week_date, y = count, color = granted_denied, fill = granted_denied)) +
  geom_bar(stat = "identity") + 
  labs(x = "", y = "Count", title = "COVID-19 Compassionate Release Cases")
ggsave(file.path(plot_path, "timeseries_all_weekly.png"), timebars_all_weekly, height = 10, width = 14)

## all - montly
timebars_all_montly <- dat %>%
  group_by(month = month(date), year = year(date), granted_denied) %>% 
  summarize(count = n()) %>%
  ungroup() %>%
  mutate(month_date = lubridate::ymd( glue("{year}-{month}-01" ))) %>%
  ggplot(., aes(x = month_date, y = count, color = granted_denied, fill = granted_denied)) +
  geom_bar(stat = "identity") + 
  labs(x = "", y = "Count", title = "COVID-19 Compassionate Release Cases")
ggsave(file.path(plot_path, "timeseries_all_monthly.png"), timebars_all_montly, height = 10, width = 14)

## denials - weekly
timebars_denial_weekly <- dat %>%
  filter(granted_denied == "denied") %>%
  group_by(week = week(date), year = year(date), granted_denied) %>% 
  summarize(count = n()) %>%
  ungroup() %>%
  # mutate(week_date = as.Date(glue('{year}-{week}-1'), "%Y-%U-%u")) %>%
  mutate(week_date = lubridate::ymd( glue("{year}-01-01" )) + lubridate::weeks(week - 1 )) %>%
  ggplot(., aes(x = week_date, y = count)) +
  geom_bar(stat = "identity") + 
  labs(x = "", y = "Count", title = "COVID-19 Compassionate Release Denials") + 
  scale_fill_manual(values=c("#D7790F")) + 
  theme_behindbars() + 
  theme(legend.title = element_blank()) 
ggsave(file.path(plot_path, "timeseries_denial_weekly.png"), timebars_denial_weekly, height = 10, width = 14)

## denials - monthly
timebars_denial_monthly <- dat %>%
  filter(granted_denied == "denied") %>%
  group_by(month = month(date), year = year(date), granted_denied) %>% 
  summarize(count = n()) %>%
  ungroup() %>%
  mutate(month_date = lubridate::ymd( glue("{year}-{month}-01" ))) %>%
  ggplot(., aes(x = month_date, y = count)) +
  geom_bar(stat = "identity", fill = "#D7790F") + 
  labs(x = "", y = "Count", title = "COVID-19 Compassionate Release Denials") + 
  geom_text(aes(label = count), vjust = -0.2) + 
  theme_behindbars() + 
  theme(legend.title = element_blank()) 
ggsave(file.path(plot_path, "timeseries_denials_monthly.png"), timebars_denial_monthly, height = 10, width = 14)

## granted - weekly 
timebars_grants_weekly <- dat %>%
  filter(granted_denied == "granted") %>%
  group_by(week = week(date), year = year(date), granted_denied) %>% 
  summarize(count = n()) %>%
  ungroup() %>%
  mutate(week_date = lubridate::ymd( glue("{year}-01-01" )) + lubridate::weeks(week - 1 )) %>%
  ggplot(., aes(x = week_date, y = count)) +
  geom_bar(stat = "identity") + 
  labs(x = "", y = "Count", title = "COVID-19 Compassionate Release Grants")
ggsave(file.path(plot_path, "timeseries_grants_weekly.png"), timebars_grants_weekly, height = 10, width = 14)

## granted - monthly 
## come back right here
timebars_grants_monthly <- dat %>%
  filter(granted_denied == "granted") %>%
  group_by(month = month(date), year = year(date), granted_denied) %>% 
  summarize(count = n()) %>%
  ungroup() %>%
  mutate(month_date = lubridate::ymd( glue("{year}-{month}-01" ))) %>%
  ggplot(., aes(x = month_date, y = count)) +
  geom_bar(stat = "identity", fill = "#4C6788") + 
  labs(x = "", y = "Count", title = "COVID-19 Compassionate Release Grants") + 
  geom_text(aes(label = count), vjust = -0.2) + 
  theme_behindbars() 
ggsave(file.path(plot_path, "timeseries_grants_monthly.png"), timebars_grants_monthly, height = 10, width = 14)

## prop granted, prop denied
n_grants <- dat %>% filter(granted_denied == "granted") %>% count() %>% pull(n)
n_denials <- dat %>% filter(granted_denied == "denied") %>% count() %>% pull(n)
timebars_prop_monthly <- dat %>%
  group_by(month = month(date), year = year(date), granted_denied) %>% 
  summarize(count = n()) %>%
  ungroup() %>%
  mutate(prop = ifelse(granted_denied == "granted", (count / n_grants), (count / n_denials)),
    month_date = lubridate::ymd( glue("{year}-{month}-01" ))) %>%
  ggplot(., aes(x = month_date, y = prop, fill = granted_denied, color = granted_denied)) +
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "", 
       y = "Status proportion (montly count / total number granted or denied) ", 
       title = "COVID-19 Compassionate Releases") 

# dat %>%
#   group_by(month = month(date), year = year(date), granted_denied) %>% 
#   summarize(count = n()) %>%
#   ungroup() %>%
#   mutate(prop = ifelse(granted_denied == "granted", (count / n_grants), (count / n_denials)),
#     month_date = lubridate::ymd( glue("{year}-{month}-01" ))) %>% 
#   ggplot(data = ., aes(month_date, prop, group = granted_denied)) +
#     geom_col(aes(fill = granted_denied), position = "dodge") +
#   ## cant get geom text right i quit!!
#     geom_text(
#       aes(label = count, 
#           y = prop + 0.0005),
#           position = position_dodge(1.8)) + 
#     labs(x = "", 
#        y = "Status proportion (montly count / total number granted or denied) ", 
#        title = "COVID-19 Compassionate Releases") + 
#   scale_y_continuous(labels = scales::percent)
ggsave(file.path(plot_path, "timeseries_prop_monthly.png"), timebars_prop_monthly, height = 10, width = 14)


## prop granted by month
monthly_count <- dat %>%
  group_by(month = month(date), year = year(date)) %>% 
  summarize(monthly_count = n()) %>%
  ungroup()

timebars_prop_by_month <- dat %>%
  group_by(month = month(date), year = year(date), granted_denied) %>% 
  summarize(count = n()) %>%
  ungroup() %>%
  mutate(granted_denied_labeller = glue('% {str_to_title(granted_denied)}')) %>%
  left_join(monthly_count) %>% 
  mutate(prop = count / monthly_count,
    month_date = lubridate::ymd( glue("{year}-{month}-01" ))) %>%
  ggplot(., aes(x = month_date, y = prop, fill = granted_denied)) +
  geom_bar(position="dodge", stat="identity", aes(fill = granted_denied)) + 
  labs(x = "", y = "Monthly % granted or denied", title = "COVID-19 Compassionate Releases")  + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + 
  scale_fill_manual(values=c("#D7790F", "#4C6788"), labels = c("% Denied", "% Granted")) + 
  theme_behindbars() + 
  theme(legend.title = element_blank()) 
ggsave(file.path(plot_path, "timeseries_prop_bymonth.png"), timebars_prop_by_month, height = 10, width = 14)
```